---
title: "Covid19 vaccine study Final2023"
author: "Colton Beardsley"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))

male = filter(FinalData, sex =="M")
female = filter(FinalData, sex =="F")
lgbtq = filter(FinalData, LGBTQ =="gay")
drugUser = filter(FinalData, DrugUser =="yes")

```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males


### Graphical Descriptive Results

```{r}
barchartGC(~infected+treatment,data=male)
barchartGC(~infected+treatment,data=male,type="percent")
```

The graph indicates a distinction of 30% between individuals with Covid-19 and those with a placebo among the males who received the vaccination and those who contracted Covid-19. In contrast, minimal disparity was observed among those who received the vaccinations and were deemed "OK."

### Numerical Descriptive Results

```{r}
table1 <- xtabs(~infected + treatment, data= male)
rowPerc(table1)
colPerc(table1)
```

The set of data illustrates a twofold difference in individuals who received the vaccinations and encountered Covid-19. Among those who took the placebo, 70% remained disease-free, in contrast to 30% of those who received the vaccination. However, there was no significant contrast among those who received vaccinations and were deemed "OK."

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


## Females


## LGBTQ



## Druggies


# Overall Results and Conclusions